Latent Predicate Networks: Concept Learning with Probabilistic Context-Sensitive Grammars

### 2014-10-21 drafts/cfp/draft.tex
final draft from Josh

\usepackage{fullpage}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{listings}
\usepackage{anyfontsize}
\usepackage{subcaption}
\usepackage{graphicx}
\usepackage{titling}
\usepackage{array}
%\usepackage{multicol}

@@

\title{Latent Predicate Networks: Concept Learning with Probabilistic Context-Sensitive Grammars}
\author{Eyal Dechter, Josh Rule, Josh Tenenbaum ADD AFFILIATION}
\date{}

\begin{document}
\vspace{-10cm}
\maketitle
\vspace{-2em}

{\small For humans, learning abstract concepts and learning language
  go hand in hand: we acquire abstract knowledge primarily through
  linguistic experience, and acquiring abstract concepts is a crucial
  step in learning the meanings of linguistic expressions. Number
  knowledge is a case in point: we largely acquire concepts such as
  seventy-three through linguistic means, and we can only know what
  the sentence ``Seventy-three is more than twice as big as
  thirty-one" means if we can grasp the meanings of its component
  number words. How do we make sense of this potential paradox? One
  approach is to estimate the distribution from which sentences are
  drawn, and, in doing so, infer the latent concepts and relationships
  that best explain those sentences. We present early work on a
  learning framework called Latent Predicate Networks (LPNs) which
  learns concepts by inferring the parameters of probabilistic
  context-sensitive grammars over sentences.  We show that for a small
  fragment of sentences expressing relationships between English
  number words, we can use hierarchical Bayesian inference to learn
  LPNs that can answer simple queries about previously unseen abstract
  concepts within this domain. These generalizations demonstrate LPNs'
  promise as a tool for learning and representing conceptual knowledge
  in language.}

@@

Although concept learning and language acquisition have typically been
treated as distinct problems in AI, linguistics and cognitive
development, they are strongly coupled for a child learning to
understand language. More generally, people learn many abstract
concepts primarily through language even though understanding language
depends on understanding the underlying concepts. Research in concept
learning is often focused on concepts grounded in perceptual features,
and while it is almost certainly true that many concepts are learned
via generalization from concrete examples, some concepts cannot
possibly be learned this way.

Number concepts are a good example: children do not learn about the
meaning of ``seventy five'' by seeing examples of seventy five things;
they do not know that ``seventy five'' is more than ``twenty five''
because of their perceptual experiences of these quantities. Rather,
children learn the meaning of ``seventy five'' (or ``a billion and
five'') by noticing how number words are used in language, in counting
sequences, in arithmetic excercises, etc. Other good examples of such
abstract concepts are kinship and social relations (e.g. ``my father
in law's grandmother''), temporal relations (``the day after last
Thanksgiving.''), and spatial relations (``above and just to the left
of'').

@@

A string $x$ is in the language defined by an RCG if one can
\emph{derive} $S(x)$. A derivation is a sequence of rewrite steps in
which substrings of the left hand side argument string are bound to
the variables of the head of some clause, thus determining the
arguments in the clause body. If a clause has no body terms, then its
head is derived; otherwise, its head is derived if its body clauses
are derived~\footnote{This description of the language of an RCG
  technically only holds for \emph{non-combinatory} RCGs, in which the
  arguments of body terms can only contain single variables. Since any
  \emph{combinatory} RCG can be converted into a non-combinatory RCG
  and we only consider non-combinatory RCGs here, this description
  suffices.}

@@

Although it is difficult to know how common these kinds of sentences
are in child-directed speech, words for small numbers are far more
common than words for larger ones(cite CHILDES). On the other hand,
children learning to count to large numbers rehearse the sequence. To
approximate this distribution of evidence, we drew these sample
sentences from a sum of a geometric distribution with parameter $0.5$
and a uniform distribution. These components were weighted $75\%$ and
$25\%$, respectively. We drew $2000$ examples from this distribution,
holding out the sentences in Table~\ref{tab:results} for
evaluation.

@@

We evaluated the learned model by asking for Viterbi (i.e. maximum a
posteriori) completions of the last words of each held out test
sentence. Table~\ref{tab:results} shows these completions. The grammar
correctly learns much of the structure of these sentences, including
the difference between sentences starting with ``before'' and
``after'' and the edge cases that relate decade words like ``twenty''
to non-decade words like ``twenty one.''

@@

Many of the algorithms and formalisms used here were originally
developed for use in other closely related areas, including logic
programming and semantic parsing. Semantic parsing, in particular,
seems especially related to the challenge we face here of jointly
learning meaning and structure from
sentences~\cite{berant2013semantic, liang2013learning,
  kwiatkowski2010inducing}. Semantic parsing, however, tends to frame
this challenge slightly differently than we do here, namely, by asking
how utterances can be mapped to an explicit internal logical
language. By contrast, we focus here on systems where meaning and
structure seem inseparable. Understanding how these two approaches
relate and inform one another is an interesting and open question.

@@

\nocite{*}

>>>>

\usepackage{fullpage}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{listings}
\usepackage{anyfontsize}
\usepackage{subcaption}
\usepackage{graphicx}
\usepackage{titling}
\usepackage{array}
\usepackage{authblk}

@@

\title{Latent Predicate Networks: Concept Learning with Probabilistic Context-Sensitive Grammars}
\author{Eyal Dechter (edechter)}
\author{Joshua Rule (rule)}
\author{Joshua B. Tenenbaum (jbt)}
\affil{Department of Brain and Cognitive Sciences, MIT (mit.edu)}

\renewcommand\Authands{ and }
\date{}

\begin{document}
\vspace{-10cm}
\maketitle
\vspace{-4em}

{\small For humans, learning abstract concepts and learning language
  go hand in hand: we acquire abstract knowledge primarily through
  linguistic experience, and acquiring abstract concepts is a crucial
  step in learning the meanings of linguistic expressions. Number
  knowledge is a case in point: we largely acquire concepts such as
  seventy-three through linguistic means, and we can only know what
  the sentence ``Seventy-three is more than twice as big as
  thirty-one" means if we can grasp the meanings of its component
  number words. How do we begin to solve this problem? One approach is
  to estimate the distribution from which sentences are drawn, and, in
  doing so, infer the latent concepts and relationships that best
  explain those sentences. We present early work on a learning
  framework called Latent Predicate Networks (LPNs) which learns
  concepts by inferring the parameters of probabilistic
  context-sensitive grammars over sentences.  We show that for a small
  fragment of sentences expressing relationships between English
  number words, we can use hierarchical Bayesian inference to learn
  LPNs that can answer simple queries about previously unseen
  relationships within this domain. These generalizations demonstrate
  LPNs' promise as a tool for learning and representing conceptual
  knowledge in language.}

@@

Although concept learning and language acquisition have typically been
treated as distinct problems in AI, linguistics and cognitive
development, they are strongly coupled for a child learning to
understand language. More generally, people learn many abstract
concepts primarily through language even though understanding language
depends on understanding the underlying concepts. Research in concept
learning is often focused on concepts grounded in perceptual features,
and while it is almost certainly true that many concepts are learned
via generalization from concrete examples, some concepts cannot be
learned this way.

Number concepts are a good example: children do not learn about the
meaning of ``seventy five'' by seeing examples of seventy five things;
they do not know that ``seventy five'' is more than ``twenty five''
because of their perceptual experiences of these quantities. Rather,
children learn the meaning of ``seventy five'' (or ``a billion and
five'') by noticing how number words are used in language, in counting
sequences, in arithmetic excercises, {\it etc}. Other good examples of such
abstract concepts are kinship and social relations ({\it e.g.} ``my father
in law's grandmother''), temporal relations (``the day after last
Thanksgiving.''), and spatial relations (``above and just to the left
of'').

@@

A string $x$ is in the language defined by an RCG if one can
\emph{derive} $S(x)$. A derivation is a sequence of rewrite steps in
which substrings of the left hand side argument string are bound to
the variables of the head of some clause, thus determining the
arguments in the clause body. If a clause has no body terms, then its
head is derived; otherwise, its head is derived if its body clauses
are derived.\footnote{This description of the language of an RCG
  technically only holds for \emph{non-combinatory} RCGs, in which the
  arguments of body terms can only contain single variables. Since any
  \emph{combinatory} RCG can be converted into a non-combinatory RCG
  and we only consider non-combinatory RCGs here, this description
  suffices.}

@@

Although it is difficult to know how common these kinds of sentences
are in child-directed speech, words for small numbers are far more
common than words for larger ones \cite{macwhinney2000childes}. On the
other hand, children learning to count to large numbers rehearse the
sequence. To approximate this distribution of evidence, we drew these
sample sentences from a sum of a geometric distribution with parameter
$0.5$ and a uniform distribution. These components were weighted
$75\%$ and $25\%$, respectively. We drew $2000$ examples from this
distribution, holding out the sentences in Table~\ref{tab:results} for
evaluation.

@@

We evaluated the learned model by asking for Viterbi ({\it i.e.} maximum a
posteriori) completions of the last words of each held out test
sentence. Table~\ref{tab:results} shows these completions. The grammar
correctly learns much of the structure of these sentences, including
the difference between sentences starting with ``before'' and
``after'' and the edge cases that relate decade words like ``twenty''
to non-decade words like ``twenty one.''

@@

Many of the algorithms and formalisms used here were originally
developed for use in other closely related areas, including logic
programming and semantic parsing. Semantic parsing, in particular,
seems especially related to the challenge we face here of jointly
learning meaning and structure from
sentences~\cite{berant2013semantic, liang2013learning,
  kwiatkowski2010inducing, poon2009unsupervised}. Semantic parsing,
however, tends to frame this challenge slightly differently than we do
here, namely, by asking how utterances can be mapped to an explicit
internal logical language. By contrast, we focus here on systems where
meaning and structure seem inseparable. Understanding how these two
approaches relate and inform one another is an interesting and open
question.

@@
